A Partitioning Algorithm for Maximum Common Subgraphs\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1 and EP/M508056/1]}

### 2017-02-15 paper/paper.tex
More writing

We initially assume that graphs are unlabelled, undirected and without self-loops
(Section~\ref{sec:extensions} describes how these restrictions may be relaxed).
The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.
The set of vertices adjacent to vertex $v$ in graph $G$ is called the
\emph{neighbourhood} of $v$, and denoted $\N(v, G)$.
We use $n_G$ and $n_H$ to denote $|\V(G)|$ and $|\V(H)|$. We will occasionally use
$n$ in complexity results to denote $n_G + n_H$.  The graph $\overline{G}$ is the complement (aka the inverse) of $G$.

@@

%At each node of the search tree, the algorithm
%labels vertices in $\V(G)$ and $\V(H)$ such that a pair $(v, w)$ with $v \in
%\V(G)$ and $w \in \V(H)$ can be added to the mapping if and only if $v$ and $w$
%have the same label.
\fi
\McSplit\ finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
(v_{|M^*|}, w_{|M^*|})\}$, where the $v_i$ are distinct members of $\V(G)$ and the
$w_i$ are distinct vertices of $\V(H)$, such that $v_i$ and $v_j$ are adjacent
in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$. It is clear that,
given such a mapping, the subgraph of $G$ induced by $\{v_i, \dots, v_{|M^*|}\}$
and the subgraph of $H$ induced by $\{w_i, \dots, w_{|M^*|}\}$ are a maximum
common subgraph.

@@

We begin by mapping vertex $1$ to vertex $a$, giving $M=\{1a\}$.  Now label
each unmatched vertex according to whether it is adjacent to vertex $1$ (in $G$) or
vertex $a$ (in $H$), as shown in Figure~\ref{fig:alg2}.  Adjacent vertices have label
1; non-adjacent vertices have label 0.  It is straightforward to see that we
can extend $M$ with a vertex mapping $vw$, with $v \in \V(G)$ and $w \in
\V(H)$, if and only if $v$ and $w$ have the same label.  This property, that
two vertices may be mapped together if and only if they share a label, is the
algorithm's main invariant.

@@

Our invariant is maintained: we can extend $M$ by a vertex pairing if
and only if the two vertices have the same label. Note also that each label (in Figure~\ref{fig:alg3}) is prefixed
by the label at the previous level of the search tree (Figure~\ref{fig:alg2}).

@@

\paragraph{Efficient representation of label classes} The algorithm as
described so far requires $\BigO{n^2}$ space at each level of the search tree,
since labels may be up to $n_G$ bits in length. We can reduce the memory
requirement to $\BigO{n}$ per level by storing a \emph{label class} as a pair $\langle P,T \rangle$ for
each label $l$ that is used, where $P$ is the set of vertices in $\V(G)$
labelled $l$, and $T$ is the set of vertices in $\V(H)$ labelled $l$\footnote{We adopt terminology from subgraph isomorphism where one graph is a pattern graph $P$ and the other the target graph $T$.}. Since there are $n_G + n_H$ vertices in the two graphs, at most $n_G + n_H$
label classes can exist at once, and the total length of the $P$ and $T$ sets
is at most $n_G + n_H$.  We can gain further efficiency by discarding during
search any label class that has an empty $P$ or $T$ set.

\begin{algorithm}[ht]
\DontPrintSemicolon
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
\nl \Begin{
%\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
%\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
\bigskip
\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
\nl \For {$w \in T$ \label{WLoop}} {
\nl    $M \gets M \cup (v,w)$ \label{GrowM} \;
\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
\nl        \bf{Let} $P'' \gets P' \cap N(v,G) \setminus \{v\}$ \label{NewPWithEdge} \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,H) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
\nl        \bf{Let} $P'' \gets P' \cap N(v,\overline{G}) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,\overline{H}) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
       }
\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
  }
\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
}
\;
\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
\nl \Begin{
\nl $\AlgVar{incumbent} \gets \emptyset$ \;
\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
\nl return $|\AlgVar{incubent}|$ \;
}
\caption{\McSplit\ algorithm for Maximum Common Subgraph}
\label{jtAlg}
\end{algorithm}

\paragraph{Algorithm \ref{jtAlg} Explained} The recursive procedure, $\FuncSty{expand}$, has
two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
each represented as a $\langle P, T \rangle$ pair as described above.  The
parameter $M$ is the current mapping of vertices.  On each call to
$\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be added to $M$
if and only if $v$ and $w$ belong to the same label class in $M$.

@@

\section{Comparison with Existing Algorithms}\label{sec:comparison}
The  \McSplit\ algorithm may be considered as a version of the
constraint programming, forward-checking (CP-FC) algorithm of \cite{DBLP:conf/cp/NdiayeS11}.
The vertices in $P$ correspond to variables in CP-FC and the 
vertices in $T$ to the domains of variables in CP-FC. A label class $\langle P,T \rangle$ then corresponds to a set 
of variables ($P$) in CP-FC  with identical domains ($T$). Where CP-FC uses a soft allDifferent constraint to filter and compute 
a bound, \McSplit\ computes a bound from each label class in the future, i.e. it performs Hall Set reasoning on the fly.

\McSplit\ takes $\BigO{n}$ time at each search node. Furthermore, \McSplit\ has a low start up cost. The clique encoding requires the
construction of the association graph prior to search

Our algorithm needs less space than CP-FC ($\BigO{n^2}$ rather than $\BigO{n^3}$?) and
much less than clique ($\BigO{n^4}$?).

@@

LARGELY FROM CIARAN'S EMAIL; NEEDS WORK:

\begin{algorithm}
\DontPrintSemicolon
\nl    \For {$q \in \N(b,G)$}{
\nl        remove $t$ from $D'_q$ \;
\nl        remove the inverse neighbourhood of $t$ from $D'_q$ \;
       }
\nl    \For {$q \in \N(b,\overline{G})$}{
\nl        remove $t$ from $D'_q$ \;
\nl        remove the neighbourhood of $t$ from $D'_q$ \;
       }
\caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} in CP algorithm}
\label{cpAlg}
\end{algorithm}

The first difference is in the underlying data structures: James uses a clever
partitioning data structure, not a conventional domain store. With the
algorithm described this way, I think the proof that domains are partitioned
into disjoint sets is almost immediate.

The second difference is in the bound function: the French code constructs a
maximum cardinality matching, with some special weighting trickery for $\bot$,
whilst James just sums the smaller of the two sides of each partition. If you
think a bit about what the matching looks like, you can see that these two give
the same answer, but at very different speeds.

The third difference is in the details of the variable ordering heuristic.

@@

\paragraph{Undirected with edge labels} Each adjacency matrix
entry contains an edge label, or a null entry (non-adjacent) $0$.
We use Algorithm~\ref{labDirAlg}, by letting $L$ be
the union of $\{0\}$ with the set of all labels that appear in the input
graphs. Since there may be up to $n_G + n_H$ distinct labels, the loop in
Algorithm~\ref{labDirAlg} may execute up to $n_G + n_H$ times and
requires $\BigO{n^2}$ time per search
node.  It is possible to modify the algorithms to run in $\BigO{n \log n}$ per
search node.  First, run lines 17-19
of Algorithm~\ref{jtAlg} to create a new label-class of vertices that are not
adjacent to $v$ or $w$, and remove these vertices from $\langle P',T' \rangle$.
Next, sort $P'$ and $T'$ in ascending order of the label on the edge from $v$
or $w$ to each vertex. We can then create the label classes corresponding to
each edge label by simultaneously traversing $P'$ and $T'$ from left to right.

@@

\iffalse
\paragraph{Almost Subgraph Isomorphism} \citet{UpcomingAAAIPaper} present a
``top-down'' algorithm for Maximum Common Subgraph in which a solution of size
$n_G$ is initially sought; if such a solution does not exist then this
solutions of size $n_G-1, n_G-2, \dots$ are sought in sequence.  This strategy,
when used with techniques based on a state-of-the-art subgraph isomorphism
solver, outperformed CP and clique encodings on a large benchmark set of graph
pairs.  We can modify the \McSplit\ algorithm to use the top-down strategy by
calling the main \FuncSty{McSplit} method once per goal size ($n_G, n_G-1,
n_G-2, \dots$).  We backtrack (\lineref{PruneSearch} of Algorithm~\ref{jtAlg})
when the bound is strictly less than the goal size, and terminate when a
solution of the goal size is found.
\fi

@@

Experiments were performed on a cluster of machines with Dual Xeon E5-2640
v2 CPUs and 64GBytes RAM. 
%Parallel experiments used 32 threads per machine,
%using all 16 physical cores with hyper-threading enabled.
All experiments used a database of
randomly-generated Maximum Common Subgraph instances, which contains graphs of
order ranging from 10 to 100 vertices
\cite{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}.  For
unlabelled instances, we selected 10 per cent of the 41,100 instances with up
to 40 vertices; labels were discarded. For labelled instances, we selected 10
per cent of the full database of 81,400 instances, and used a labelling scheme
in which the number of distinct vertex labels and the number of distinct edge
labels is approximately equal to 33 per cent of the number of vertices in each
graph.

@@

\begin{figure}[ht]
    \centering
    \subfigure[][MCS unlabelled undirected instances] {
        \centering
        \includegraphics*{gen-graph-plain-cumulative.pdf}
        \label{figure:plain-cumulative}
    }

@@

\begin{figure}
    \centering
    \subfigure[][MCS unlabelled undirected instances] {
        \centering
        \includegraphics*{gen-graph-plain-connected-cumulative.pdf}
        \label{figure:plain-connected-cumulative}
    }

>>>>

We initially assume that graphs are unlabelled, undirected and without loops
(Section~\ref{sec:extensions} describes how these restrictions may be relaxed).
The vertices and edge sets of a graph $G$ are denoted $\V(G)$ and $\E(G)$.  The
set of vertices adjacent to vertex $v$ in graph $G$ is called the
\emph{neighbourhood} of $v$, and denoted $\N(v, G)$.  The graph $\overline{G}$
is the complement of $G$, obtained copying the vertices of $G$ and adding an
edge between each pair of distinct vertices if and only if $G$ does not
have an edge between the corresponding pair.

Let $G$ and $H$ be the two input graphs to our maximum common subgraph problem.
The orders (vertex counts) of these graphs are denoted $n_G$ and $n_H$. We will
use $n$ in complexity results to denote $\max(n_G + n_H)$.

@@

%At each node of the search tree, the algorithm
%labels vertices in $\V(G)$ and $\V(H)$ such that a pair $(v, w)$ with $v \in
%\V(G)$ and $w \in \V(H)$ can be added to the mapping if and only if $v$ and $w$
%have the same label.
\fi
\McSplit\ finds a maximum-cardinality mapping $M^* = \{(v_1, w_1), \dots,
(v_{|M^*|}, w_{|M^*|})\}$, where the $v_i$ are distinct members of $\V(G)$ and the
$w_i$ are distinct vertices of $\V(H)$, such that $v_i$ and $v_j$ are adjacent
in $G$ if and only if $w_i$ and $w_j$ are adjacent in $H$. 
Given such a mapping, the subgraph of $G$ induced by $\{v_i, \dots, v_{|M^*|}\}$
and the subgraph of $H$ induced by $\{w_i, \dots, w_{|M^*|}\}$ are a maximum
common subgraph.

@@

We begin by mapping vertex $1$ to vertex $a$, giving $M=\{1a\}$.  Now label
each unmatched vertex according to whether it is adjacent to vertex $1$ (in
$G$) or vertex $a$ (in $H$), as shown in Figure~\ref{fig:alg2}.  Adjacent
vertices have label 1; non-adjacent vertices have label 0.  We can extend $M$
with a vertex mapping $vw$, with $v \in \V(G)$ and $w \in \V(H)$, if and only
if $v$ and $w$ have the same label.  This property, that two vertices may be
mapped together if and only if they share a label, is the algorithm's main
invariant.

@@

Our invariant is maintained: we can extend $M$ by a vertex pairing if and only
if the two vertices have the same label. Note also that the labels can be
calculated efficiently, since each label (in Figure~\ref{fig:alg3}) is prefixed
by the label at the previous level of the search tree (Figure~\ref{fig:alg2}).

@@

\paragraph{Efficient representation of label classes} The algorithm as
described so far requires $\BigO{n^2}$ space at each level of the search tree,
since labels may be up to $n_G$ bits in length. We can reduce the memory
requirement to $\BigO{n}$ per level by storing a \emph{label class} as a pair
$\langle P,T \rangle$ for each label $l$ that is used, where $P$ is the set of
vertices in $\V(G)$ labelled $l$, and $T$ is the set of vertices in $\V(H)$
labelled $l$.\footnote{We adopt terminology from subgraph isomorphism where one
graph is a pattern graph $P$ and the other the target graph $T$.} Since there
are $n_G + n_H$ vertices in the two graphs, at most $n_G + n_H$ label classes
can exist at once, and there are at most $n_G + n_H$ vertices in the union of
all of the $P$ and $T$ sets.  We can gain further efficiency by discarding
during search any label class that has an empty $P$ or $T$ set.

\begin{algorithm}[ht]
\DontPrintSemicolon
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
\nl \Begin{
%\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
%\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
\bigskip
\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
\nl \For {$w \in T$ \label{WLoop}} {
\nl    $M \gets M \cup (v,w)$ \label{GrowM} \;
\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
\nl        \bf{Let} $P'' \gets P' \cap N(v,G) \setminus \{v\}$ \label{NewPWithEdge} \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,H) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
\nl        \bf{Let} $P'' \gets P' \cap N(v,\overline{G}) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
\nl        \bf{Let} $T'' \gets T' \cap N(w,\overline{H}) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}
        {$\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
       }
\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
  }
\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
}
\;
\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
\nl \Begin{
\nl $\AlgVar{incumbent} \gets \emptyset$ \;
\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
\nl return $|\AlgVar{incubent}|$ \;
}
\caption{\McSplit\ algorithm for Maximum Common Subgraph}
\label{McCliqueAlg}
\end{algorithm}

\paragraph{Algorithm \ref{McCliqueAlg} Explained} The recursive procedure, $\FuncSty{expand}$, has
two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
each represented as a $\langle P, T \rangle$ pair as described above.  The
parameter $M$ is the current mapping of vertices.  On each call to
$\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be added to $M$
if and only if $v$ and $w$ belong to the same label class in $M$.

@@

\section{Comparison with Existing Algorithms}\label{sec:comparison} The
\McSplit\ algorithm may be considered as a version of the constraint
programming, forward-checking (CP-FC) algorithm of
\citep{DBLP:conf/cp/NdiayeS11}, using an efficient, compressed representation
of domains.  (Recall that in the CP-FC algorithm, each vertex in $v \in \V(G)$
is represented by a variable, whose domain corresponds to the set of vertices
in $\V(H)$ to which $v$ may currently be mapped.)

Given a label class $\langle P,T \rangle$ in \McSplit\, the vertices in $P$
correspond to variables in CP-FC, all of which have domain equal to $T$.  The
label-class representation of domains is possible because throughout the CP-FC
algorithm for max common subgraph, the domains of any two variables are either
identical or disjoint.  (To the best of our knowledge, this observation has
not been made previously.)

Where CP-FC uses a soft allDifferent constraint to filter and compute a bound,
\McSplit\ computes a bound from each label class in the future, i.e. it
performs Hall Set reasoning on the fly.

\McSplit\ requires $\BigO{n}$ time at each search node, and does not require
any auxiliary data structures larger than the adjacency matrices of $G$ and $H$
to be built prior to search.  CP-FC requires $\BigO(?????)$ \dots The clique
encoding requires an association graph with $n_G n_H$ vertices to be
constructed before search.  Modern maximum-clique algorithms typically perform
a colouring step at each search node that is quadratic in the order of the
graph, and thus $\BigO{n^4}$ in the order of the max common subgraph input
graphs.

@@

The correctness of \McSplit\ can be seen by noting its equivalence to CP-FC
\citep{DBLP:conf/cp/NdiayeS11}, if we replace the label-class representation of
$\mathit{future}$ with domain stores in Algorithm \ref{McCliqueAlg}.  To obtain
the CP-FC algorithm, replace \linerangeref{InnerLoop}{InnerLoopEnd} of
Algorithm \ref{McCliqueAlg} with Algorithm \ref{cpAlg}.  The pruning of domains
carried out in these lines can be easily seen to be equivalent to the
corresponding lines in Algorithm \ref{McCliqueAlg}.

\begin{algorithm}
\DontPrintSemicolon
\nl    \For {$u \in \N(b,G)$}{
\nl        remove $v$ from $D'_u$ \;
\nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
       }
\nl    \For {$u \in \N(b,\overline{G})$}{
\nl        remove $v$ from $D'_u$ \;
\nl        remove the neighbourhood of $t$ from $D'_u$ \;
       }
\caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm}
\label{cpAlg}
\end{algorithm}

TO WRITE PROPERLY: The second difference is in the bound function: the French
code constructs a maximum cardinality matching, with some special weighting
trickery for $\bot$, whilst James just sums the smaller of the two sides of
each partition. If you think a bit about what the matching looks like, you can
see that these two give the same answer, but at very different speeds.

@@

\paragraph{Undirected with edge labels} Each adjacency matrix
entry contains an edge label, or a null entry (non-adjacent) $0$.
We use Algorithm~\ref{labDirAlg}, by letting $L$ be
the union of $\{0\}$ with the set of all labels that appear in the input
graphs. Since there may be up to $n_G + n_H$ distinct labels, the loop in
Algorithm~\ref{labDirAlg} may execute up to $n_G + n_H$ times and
requires $\BigO{n^2}$ time per search
node.  It is possible to modify the algorithms to run in $\BigO{n \log n}$ per
search node.  First, run lines 17-19
of Algorithm~\ref{McCliqueAlg} to create a new label-class of vertices that are not
adjacent to $v$ or $w$, and remove these vertices from $\langle P',T' \rangle$.
Next, sort $P'$ and $T'$ in ascending order of the label on the edge from $v$
or $w$ to each vertex. We can then create the label classes corresponding to
each edge label by simultaneously traversing $P'$ and $T'$ from left to right.

@@

\paragraph{Top-Down Strategy} I'VE RE-ADDED A SHORTER VERSION OF THIS
PARAGRAPH, WHICH WAS COMMENTED OUT BY PATRICK. IT NEEDS WORK, BUT WE NEED TO
EXPLAIN THE SIP EXPERIMENT. We can modify the \McSplit\ algorithm to use a
top-down strategy based on \citet{UpcomingAAAIPaper} by calling the main
\FuncSty{McSplit} method once per goal size ($n_G, n_G-1, n_G-2, \dots$).  We
backtrack (\lineref{PruneSearch} of Algorithm~\ref{McCliqueAlg}) when the bound is
strictly less than the goal size, and terminate when a solution of the goal
size is found.

@@

Experiments were performed on a cluster of machines with Dual Xeon E5-2640
v2 CPUs and 64GBytes RAM.

? MENTION PROGRAMMING LANGUAGES, COMPILER FLAGS, MCSa1 ALGORITHM

%Parallel experiments used 32 threads per machine,
%using all 16 physical cores with hyper-threading enabled.
All experiments used a database of
randomly-generated Maximum Common Subgraph instances, which contains graphs of
order ranging from 10 to 100 vertices
\cite{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}.  For
unlabelled instances, we selected 10 per cent of the 41,100 instances with up
to 40 vertices; labels were discarded. For labelled instances, we selected 10
per cent of the full database of 81,400 instances, and used a labelling scheme
in which the number of distinct vertex labels and the number of distinct edge
labels is approximately equal to 33 per cent of the number of vertices in each
graph.

@@

\begin{figure}[h]
    \centering
    \subfigure[][MCS unlabelled undirected instances] {
        \centering
        \includegraphics*{gen-graph-plain-cumulative.pdf}
        \label{figure:plain-cumulative}
    }

@@

\begin{figure}[ht]
    \centering
    \subfigure[][MCS unlabelled undirected instances] {
        \centering
        \includegraphics*{gen-graph-plain-connected-cumulative.pdf}
        \label{figure:plain-connected-cumulative}
    }
